---
title: "Time Series Study on CO2 Emissions and Hybrid Car Sales Project"
author: "Grace Lang & Jules Stacy"
date: "3/18/2021"
output: html_document
---
# Data sources: 
* Hybrid Car sales: https://afdc.energy.gov/data/10301
* CO2 Emissions by Country: https://data.worldbank.org/indicator/EN.ATM.CO2E.PC?view=map

* Turbine data citation:
* Hoen, B.D., Diffendorfer, J.E., Rand, J.T., Kramer, L.A., Garrity, C.P., and Hunt, H.E., 2018, United States Wind Turbine Database (ver. 3.3, January 14, 2021): U.S. Geological Survey, American Clean Power Association, and Lawrence Berkeley National Laboratory data release, https://doi.org/10.5066/F7TX3DN0.

# EDA / Notes for Initial Presentation: 
a.	Identify yourself and your partner (if applicable).  Goal 1
b.	Describe the Data Set / Time Series (Who, What, When, Where, Why and How) Goal 1
c.	Stationary / Non-Stationary Goal 2
d.	ACFs and Spectral Densities just to explore Goal 2
e.	At least 2 (total) candidate ARMA / ARIMA models (univariate) Goal 2
    a.	The models in factored form or at least separate the stationary and non-
stationary factors with standard deviation or variance of the white noise.
    b.	AIC
    c.	ASE (Rolling window)
    d.	Visualization of Forecasts (Short and Long Term / State and National) 
f.	Strategy / Plans for the rest of the analysis.


```{r}
# install.packages('PerformanceAnalytics')

```



```{r setup, include=FALSE}
library(tswge)
library(dplyr)
library(DataCombine)
library(PerformanceAnalytics)
```


```{r}


#rolling window ASE
R_W_ASE = function(series, trainingSize, horizon=1, s=0, d=0, phis=0, thetas=0)
{
  trainingSize=trainingSize
  horizon=horizon
  ASEHolder=numeric()
  s=s
  d=d
  phis=phis
  thetas=thetas

for( i in 1:(length(series)-(trainingSize + horizon) + 1))
{
  forecasts = fore.aruma.wge(series[i:(i+(trainingSize-1))], phi=phis, theta=thetas, s=s, d=d, n.ahead=horizon)

  ASE=mean((series[(trainingSize+i):(trainingSize + i + (horizon) - 1)] - forecasts$f)^2)
  print(paste0('i: ', i))
  print(paste0('ASE: ', ASE))
  ASEHolder[i] = ASE
}
  ASEHolder
  hist(ASEHolder)
  WindowedASE = mean(ASEHolder)

  print("The Summary Statistics for the Rolling Window ASE Are:")
  print(summary(ASEHolder))
  print(paste("The Rolling Window ASE is: ", WindowedASE))
  return(WindowedASE)
}




```


```{r}

car_orig <- read.csv("./HybridCar.csv")

#Remove years that are majority blank
car <- car_orig[40:60,]

#Rename column to something easier
names(car)[names(car) == "CO2.emissions..metric.tons.per.capita."] <- "CO2"

#Loading in the wind turbine data
turbine <- read.csv("./turbines.csv")

turbine_tot = aggregate(turbine$p_tnum, by=list(Category=turbine$p_year), FUN=sum)
turbine_tot
#Renaming wind turbine column
names(turbine_tot)[names(turbine_tot) == "x"] <- "wind"
plotts.sample.wge(turbine_tot$wind)

#Jules - this merge didnt pull in the wind turbine data 
#alldat = merge(car_orig, turbine_tot, by.x=c('USA'), by.y=c('Category'), all=TRUE)


# Data cleanup

#turbine data is missing 1993
x<- turbine_tot[1:35,]
x<- x %>% add_row(Category = 1993, wind = 0, .before = 13)
#shorten to just the years with cross-over data
y<- car_orig[22:57,1:3]
#combine the 2 datasets
alldata <- cbind(x,y)
#alldata <- select(alldata,-USA) #remove duplicate year column that was for reference
#Rename column to something easier
names(alldata)[names(alldata) == "CO2.emissions..metric.tons.per.capita."] <- "CO2"

```

# Initial look at data plotted out
Data cut differently by years where data is available





```{r}
plotts.wge(car$CO2)
plotts.wge(car$TotalSales)

car_1 <- car_orig[40:57,]
plotts.sample.wge(car_1$CO2)

co2_total <- car_orig[1:57,]
names(co2_total)[names(co2_total) == "CO2.emissions..metric.tons.per.capita."] <- "CO2"
plotts.sample.wge(co2_total$CO2)

length_CO2 = length(co2_total$CO2)
```



# First take of some ARIMA Models
```{r}
#ARIMA Model Plotting
plotts.sample.wge(co2_total$CO2)
aic5.wge(co2_total$CO2,p=0:10) #AIC shows an ARMA(1,1) s best fit 
aic5.wge(co2_total$CO2, type="bic") #BIC shows an ARMA(1,1) s best fit 

#fit the model -- ARMA(1.1)
est.arma.wge(co2_total$CO2, p=1, q=1)
m1 = fore.arma.wge(co2_total$CO2,phi=0.9261945,theta = -0.5741199 , n.ahead = 5,lastn=TRUE,  plot=TRUE)
mean((co2_total$CO2[53:57]-m1$f)^2)
#ASE =  0.8390832

# #Using the ARMA(1,1) to plot the 5 years ahead
# fore.arma.wge(co2_total$CO2,phi=0.9261945,theta = -0.5741199 , n.ahead = 5,lastn=FALSE,  plot=TRUE)
# #Using the ARMA(1,1) to plot the 20 years ahead
# fore.arma.wge(co2_total$CO2,phi=0.9261945,theta = -0.5741199 , n.ahead = 20,lastn=FALSE,  plot=TRUE)


#ASE's and Forecasts ================
#5 years out
n.ahead=5
CO2_f = fore.aruma.wge(x=co2_total$CO2, phi=0.9261945, theta=-0.5741199, n.ahead=n.ahead, lastn=TRUE)
#calculate ASE
ASE_ar11 = mean((CO2_f$f - co2_total$CO2[(length_CO2-n.ahead):length_CO2])^2)
#display ASE
print(paste0("5 year: ", ASE_ar11))

#forecast
fore.aruma.wge(x=co2_total$CO2, phi=0.9261945, theta=-0.5741199, n.ahead=n.ahead, lastn=FALSE)

#20 years out
n.ahead=20
CO2_f = fore.aruma.wge(x=co2_total$CO2, phi=0.9261945, theta=-0.5741199, n.ahead=n.ahead, lastn=TRUE)
#calculate ASE
ASE_ar11 = mean((CO2_f$f - co2_total$CO2[(length_CO2-n.ahead):length_CO2])^2)
#display ASE
print(paste0("20 year: ", ASE_ar11))

#forecast
fore.aruma.wge(x=co2_total$CO2, phi=0.9261945, theta=-0.5741199, n.ahead=n.ahead, lastn=FALSE)



#rolling window ASE
R_W_ASE(co2_total$CO2, phis=0.9261945, thetas=-0.5741199, trainingSize=20, horizon=5)
```



```{r}
#remove a D term

factor.wge(phi=c(rep(0, 9), 1))
est.arma.wge(co2_total$CO2, p=14)

#differencing a d=1 seasonal component
co2_diff = artrans.wge(co2_total$CO2, phi.tr=c(1))
#fit the differenced data
aic5.wge(co2_diff) #AR(0,1) recommended
co2_diff_est = est.arma.wge(co2_diff, 0, 1) #est AR(0,1) on differenced data
co2_diff_est
#check residuals
ljung.wge(co2_diff_est$res, K=24) #p = 0.940, > 0.05
ljung.wge(co2_diff_est$res, K=48) #p = 0.303, > 0.05
#residuals appear to be whitened



#ASE's and Forecasts
#5 years out
n.ahead=5
CO2_f = fore.aruma.wge(x=co2_total$CO2, phi=0, theta=c(-0.5532092), d=1, s=0, n.ahead=n.ahead, lastn=TRUE)
#calculate ASE
ASE_ar011 = mean((CO2_f$f - co2_total$CO2[(length_CO2-n.ahead):length_CO2])^2)
#display ASE
ASE_ar011

#forecast
fore.aruma.wge(x=co2_total$CO2, phi=0, theta=c(-0.5532092), d=1, s=0, n.ahead=n.ahead, lastn=FALSE)

#20 years out
n.ahead=20
CO2_f = fore.aruma.wge(x=co2_total$CO2, phi=0, theta=c(-0.5532092), d=1, s=0, n.ahead=n.ahead, lastn=TRUE)
#calculate ASE
ASE_ar011 = mean((CO2_f$f - co2_total$CO2[(length_CO2-n.ahead):length_CO2])^2)
#display ASE
ASE_ar011

#forecast
fore.aruma.wge(x=co2_total$CO2, phi=0, theta=c(-0.5532092), d=1, s=0, n.ahead=n.ahead, lastn=FALSE)



#rolling window ASE
R_W_ASE(co2_total$CO2, trainingSize=20, horizon=5, s=0, d=1, thetas=c(-0.5532092), phis=0)
```

