---
title: "Time Series Study on CO2 Emissions and Hybrid Car Sales Project"
author: "Grace Lang & Jules Stacy"
date: "3/18/2021"
output: html_document
---

# Business Objective: 
Taking care of our environment now will help our health and the health of future generations. Air pollution from vehicle CO2 Emissions can cause respiratory diseases and cancer. 

In addition to emitting less CO2, hybrid cars have had sales on the rise in recent years due to their fuel efficiency and environmental benefits.

Popularity of wind turbines have also increased over time as more landowners are willing to get them installed on their land for energy rights and economic benefit. 
Our mission is to identify and provide insights of macro trends within the environmental industry. 



# Data sources: 
* Hybrid Car sales: https://afdc.energy.gov/data/10301
* CO2 Emissions by Country: https://data.worldbank.org/indicator/EN.ATM.CO2E.PC?view=map
* Turbine data citation: https://doi.org/10.5066/F7TX3DN0


```{r}
# install.packages('PerformanceAnalytics')
```


```{r setup, include=FALSE}
## Packages to load:
library(tswge)
library(dplyr)
library(DataCombine)
library(PerformanceAnalytics)
library(vars)
library(nnfor)
```

## Rolling Window ASE Function (ARIMA):
This function calculates the ASE along a rolling window that iterates over the data from start to finish.

```{r}
R_W_ASE = function(series, trainingSize, horizon=1, s=0, d=0, phis=0, thetas=0)
{
  trainingSize=trainingSize
  horizon=horizon
  ASEHolder=numeric()
  s=s
  d=d
  phis=phis
  thetas=thetas

for( i in 1:(length(series)-(trainingSize + horizon) + 1))
{
  forecasts = fore.aruma.wge(series[i:(i+(trainingSize-1))], phi=phis, theta=thetas, s=s, d=d, n.ahead=horizon, plot=FALSE)

  ASE=mean((series[(trainingSize+i):(trainingSize + i + (horizon) - 1)] - forecasts$f)^2)
  print(paste0('i: ', i))
  print(paste0('ASE: ', ASE))
  ASEHolder[i] = ASE
}
  ASEHolder
  hist(ASEHolder)
  WindowedASE = mean(ASEHolder)

  print("The Summary Statistics for the Rolling Window ASE Are:")
  print(summary(ASEHolder))
  print(paste("The Rolling Window ASE is: ", WindowedASE))
  return(WindowedASE)
}
```

#Rolling Window ASE function (VAR)

```{r}
R_W_ASE_VAR = function(series, trainingSize, horizon=1, model)
{
  trainingSize=trainingSize
  horizon=horizon
  model=model #have to pass in the VAR model
  ASEHolder=numeric()


for( i in 1:(length(series)-(trainingSize + horizon) + 1))
{

  #forecasts = fore.aruma.wge(series[i:(i+(trainingSize-1))], phi=phis, theta=thetas, s=s, d=d, n.ahead=horizon, plot=FALSE)

  #develop a forecast given the chosen VAR model
  forecasts = predict(model,n.ahead = horizon)
  
  ASE=mean((series[(trainingSize+i):(trainingSize + i + (horizon) - 1)] - forecasts$fcst$CO2[,1])^2)
  print(paste0('i: ', i))
  print(paste0('ASE: ', ASE))
  ASEHolder[i] = ASE
}
  ASEHolder
  hist(ASEHolder)
  WindowedASE = mean(ASEHolder)

  print("The Summary Statistics for the Rolling Window ASE Are:")
  print(summary(ASEHolder))
  print(paste("The Rolling Window ASE is: ", WindowedASE))
  return(WindowedASE)
}


```

# Rolling Window ASE Function (Ensemble)
```{r}
R_W_ASE_ENS = function(series, trainingSize, horizon=1, varmodel, s=0, d=0, phis=0, thetas=0)
{
  trainingSize=trainingSize
  horizon=horizon
  varmodel=varmodel #have to pass in the VAR model
  ASEHolder=numeric()


for( i in 1:(length(series)-(trainingSize + horizon) + 1))
{

  #forecasts = fore.aruma.wge(series[i:(i+(trainingSize-1))], phi=phis, theta=thetas, s=s, d=d, n.ahead=horizon, plot=FALSE)

  #develop a forecast given the chosen VAR model
  armafore = fore.aruma.wge(series[i:(i+(trainingSize-1))], phi=phis, theta=thetas, s=s, d=d, n.ahead=horizon, plot=FALSE)
  
  forecasts = predict(varmodel,n.ahead = horizon)
  forecasts = (forecasts$fcst$CO2[,1] + armafore$f)/2
  
  ASE=mean((series[(trainingSize+i):(trainingSize + i + (horizon) - 1)] - forecasts)^2)
  print(paste0('i: ', i))
  print(paste0('ASE: ', ASE))
  ASEHolder[i] = ASE
}
  ASEHolder
  hist(ASEHolder)
  WindowedASE = mean(ASEHolder)

  print("The Summary Statistics for the Rolling Window ASE Are:")
  print(summary(ASEHolder))
  print(paste("The Rolling Window ASE is: ", WindowedASE))
  return(WindowedASE)
}


```

#Static end-of-data ASE function
```{r}
#ASE's and Forecasts ================
#5 years out
#n.ahead=5
#series = alldata$CO2
StaticASE = function(series, forecast, n.ahead){
  series = series
  forecast = forecast
  n.ahead = n.ahead

ASE = mean((forecast - series[(length_all-n.ahead+1):length_all])^2)
#display ASE: 0.501437
print(paste0(n.ahead, " year ASE: ", ASE))
}

```


# Loading in data & cleanup:
```{r}
#Loading in the CO2 Emission data
car_orig <- read.csv("./HybridCar.csv")
#Remove years that are majority blank
car <- car_orig[40:60,] #This may not be used in the modeling

#Rename column to something easier
names(car)[names(car) == "CO2.emissions..metric.tons.per.capita."] <- "CO2"

#Loading in the wind turbine data
turbine <- read.csv("./turbines.csv")
turbine_tot = aggregate(turbine$p_tnum, by=list(Category=turbine$p_year), FUN=sum)
turbine_tot
#Renaming wind turbine column
names(turbine_tot)[names(turbine_tot) == "x"] <- "wind"
plotts.sample.wge(turbine_tot$wind)


# Data cleanup

#turbine data is missing 1993
x<- turbine_tot[1:35,]
x<- x %>% add_row(Category = 1993, wind = 0, .before = 13)
#shorten to just the years with cross-over data
y<- car_orig[22:57,1:3]
#combine the 2 datasets
alldata <- cbind(x,y)
#alldata <- select(alldata,-USA) #remove duplicate year column that was for reference
#Rename column to something easier
names(alldata)[names(alldata) == "CO2.emissions..metric.tons.per.capita."] <- "CO2"

# A selection of the data from the 3 datasets will be used for modeling. 
  #All NA's will be replaced with 0
alldata[is.na(alldata)] = 0
length_all = length(alldata$CO2)

#building dataframes for later use
last5 = alldata[(length_all-4):length_all,]
minuslast5 = alldata[0:(length_all-5),]

last20 = alldata[(length_all-19):length_all,]
minuslast20 = alldata[0:(length_all-20),]
```


# Initial look at data plotted out
Data cut differently by years to see if we need to subset the data for a specific timeframe. 
```{r}
plotts.wge(alldata$CO2)
plotts.wge(alldata$TotalSales)

car_1 <- car_orig[40:57,]
plotts.sample.wge(car_1$CO2)

plotts.sample.wge(alldata$CO2)

```


# ARMA Model #1:
```{r}
#ARIMA Model Plotting
plotts.sample.wge(alldata$CO2)
aic5.wge(alldata$CO2,p=0:10) #AIC shows an ARMA(2,1) is best fit 
aic5.wge(alldata$CO2, type="bic") #BIC shows an ARMA(1,1) is best fit 

#fit the model -- ARMA(1.1)
est.arma.wge(alldata$CO2, p=1, q=1)
m1 = fore.arma.wge(alldata$CO2,phi=0.9126689,theta = -0.5966692 , n.ahead = 5,lastn=TRUE,  plot=TRUE)
#mean((alldata$CO2[53:57]-m1$f)^2)
#ASE =  0.8390832

# #Using the ARMA(1,1) to plot the 5 years ahead
# fore.arma.wge(co2_total$CO2,phi=0.9261945,theta = -0.5741199 , n.ahead = 5,lastn=FALSE,  plot=TRUE)
# #Using the ARMA(1,1) to plot the 20 years ahead
# fore.arma.wge(co2_total$CO2,phi=0.9261945,theta = -0.5741199 , n.ahead = 20,lastn=FALSE,  plot=TRUE)


#ASE's and Forecasts ================
#5 years out
n.ahead=5
CO2_f5 = fore.aruma.wge(x=alldata$CO2, phi=0.9261945, theta=-0.5741199, n.ahead=n.ahead, lastn=TRUE)

#calculate ASE
StaticASE(series=alldata$CO2, forecast=CO2_f5$f, n.ahead=n.ahead) 
#"5 year ASE: 0.740201800887329" 

#forecast
armafore5=fore.aruma.wge(x=alldata$CO2, phi=0.9261945, theta=-0.5741199, n.ahead=n.ahead, lastn=FALSE) #also used in ensemble


#20 years out
n.ahead=20
CO2_f20 = fore.aruma.wge(x=alldata$CO2, phi=0.9261945, theta=-0.5741199, n.ahead=n.ahead, lastn=TRUE)
#calculate ASE
StaticASE(series=alldata$CO2, forecast=CO2_f5$f, n.ahead=n.ahead) 
#"20 year ASE: 4.84096667545707" 

#forecast
armafore20 = fore.aruma.wge(x=alldata$CO2, phi=0.9261945, theta=-0.5741199, n.ahead=n.ahead, lastn=FALSE) #also used in ensemble



#rolling window ASE
#R_W_ASE(alldata$CO2, phis=0.9261945, thetas=-0.5741199, trainingSize=20, horizon=5)
#"The Rolling Window ASE is:  1.60880823919367"
```


# ARIMA Model #2:
An arima model will be fit to the data. Some assumptions for this model will be that the data is non-stationary and non-seasonal, and that the data is correlation based rather than signal-plus-noise based. This is reinforced by roots close to 1 in an over-fit factor table.
```{r}
#Look at the factor table
est.arma.wge(alldata$CO2, p=8)

#differencing a d=1 seasonal component
co2_diff = artrans.wge(alldata$CO2, phi.tr=c(1))
#fit the differenced data
aic5.wge(co2_diff) #AR(1,1) recommended
co2_diff_est = est.arma.wge(co2_diff, 1, 1) #est AR(1,1) on differenced data
co2_diff_est
#check residuals
ljung.wge(co2_diff_est$res, K=24) #p = 0.892, > 0.05
ljung.wge(co2_diff_est$res, K=34) #p = 0.45, > 0.05
#residuals appear to be whitened
#second Ljung test length is 34 due to data length


#ASE's and Forecasts ======================
#5 years out
n.ahead=5
CO2_f = fore.aruma.wge(x=alldata$CO2, phi=c(-0.4536403), theta=c(-0.8277382), d=1, s=0, n.ahead=n.ahead, lastn=TRUE)
#calculate ASE
StaticASE(series=alldata$CO2, forecast=CO2_f$f, n.ahead=n.ahead) 
#"5 year ASE: 0.374575891938278" 

#forecast
fore.aruma.wge(x=alldata$CO2, phi=c(-0.4536403), theta=c(-0.8277382), d=1, s=0, n.ahead=n.ahead, lastn=FALSE)



#20 years out
n.ahead=20
CO2_f = fore.aruma.wge(x=alldata$CO2, phi=c(-0.4536403), theta=c(-0.8277382), d=1, s=0, n.ahead=n.ahead, lastn=TRUE)
#calculate ASE
StaticASE(series=alldata$CO2, forecast=CO2_f$f, n.ahead=n.ahead) 
#"20 year ASE: 3.97997954166426"

#forecast
fore.aruma.wge(x=alldata$CO2, phi=c(-0.4536403), theta=c(-0.8277382), d=1, s=0, n.ahead=n.ahead, lastn=FALSE)



#rolling window ASE
#R_W_ASE(co2_total$CO2, trainingSize=20, horizon=5, s=0, d=1, thetas=c(-0.5532092), phis=c(-0.4536403))
#"The Rolling Window ASE is:  0.992686708185353"
```

# VAR Model with additions of the wind turbine & hybrid car sales data:

## Question for Sadler - i run the ASE off of past data, but then i have to reforecast with it looking ahead?


The VARselect method does not like working with the data, and suggests that AIC is negative infinity for lags above 7. When modeled at lag 7, the forecasts dropped to values far below zero, at magnitudes of eight times the maximal value in the dataset. This seems unreasonable, so a lower lag value will be selected.
```{r}
#Looking to see if there is any cross correlation with the 2 additional variables
#running(alldata$CO2, alldata$wind, fun=cor)
ccf( alldata$wind,alldata$CO2,main="CO2 Emissions to Wind Turbines")
#There is strong correlation between Wind Turbines being a leading indicator of CO2 emissions from 0-4 years.
ccf(alldata$TotalSales,alldata$CO2,main="CO2 Emissions to Hybrid Car Sales")
#There is strong correlation that it takes 0 - 8 years for Hybrid car sales to make any changes on CO2 sales

#================================
#combine data into single dataframe
df_all3 = cbind(CO2=alldata$CO2, wind=alldata$wind, cars=alldata$TotalSales)

#VARS
VARselect(df_all3, type = "both", lag.max = 10)



#VARselect picks p=9, but this is due to AIC and BIC of negative infinity. The last lag with feasible values for all metrics is lag 7.
varmodel = VAR(df_all3, p=3, type = "both")

AIC(varmodel) #1296.173

fcst5yr=predict(varmodel, n.ahead=5)$fcst$CO2[,1]



#ASE Calculations =========================================
#----------------------------
#5 year ASE calculations
#----------------------------

#Variable Prep
#forecast using last n.ahead as comparison points
#combine data into single dataframe
all3_5 = cbind(CO2=minuslast5$CO2, wind=minuslast5$wind, cars=minuslast5$TotalSales)
var5 = VAR(all3_5, p=3, type = "both")
preds5 = predict(var5, n.ahead = 5)
varase5 = preds5$fcst$CO2[,1]

#Plotting
#plot the forecast of the last 5 points
plot(seq(1, length(alldata[,1]), 1), alldata$CO2, type='b')
#the plot thickens (adding prediction points)
points(seq((length(alldata[,1])-4), length(alldata[,1]), 1), varase5, type='b', pch=5)

#Forecasting
#5 year 
StaticASE(series=alldata$CO2, forecast=varase5, n.ahead=5) 
#"5 year ASE: 3.67002304400658"


#20 year ASE calculations ---------------
#unable to use cars because there are no sales data for the first 16 years
all3_20 = cbind(CO2=minuslast20$CO2, wind=minuslast20$wind) 
var20 = VAR(all3_20, p=3, type = "both")
preds20 = predict(var20, n.ahead = 20)
varase20 = preds20$fcst$CO2[,1]

#plot the forecast of the last 20 points
plot(seq(1, length(alldata[,1]), 1), alldata$CO2, type='b')
#the plot thickens (adding prediction points)
points(seq((length(alldata[,1])-19), length(alldata[,1]), 1), varase20, type='b', pch=5)

all3_20

# 20 year
StaticASE(series=alldata$CO2, forecast=varase20, n.ahead=20) 
#"20 year ASE: 2.61548951244489" 
  
  
#Forecasts ============================
#5 year ahead
n.ahead=5
preds5 = predict(varmodel,n.ahead = n.ahead) #forecast the next 5 years 
varfore5 = preds5$fcst$CO2[,1] #also used in ensemble modeling

#20 year ahead
n.ahead=20
preds20 = predict(varmodel,n.ahead = n.ahead) #forecast the next 20 years
varfore20 = preds20$fcst$CO2[,1] #used in ensemble modeling

#Plotting =============================
#Plotting out the 5 Year forecast
plot(seq(1,36,1),alldata$CO2,type="b", xlim=c(1,41), ylab="CO2 Emissions",main="VAR - 5 Year Forecast")
points(seq(37,41,1),varfore5,type="b",pch=15)

#Plotting out the 20 Year forecast
plot(seq(1,36,1),alldata$CO2,type="b", xlim=c(1,56), ylim =c(12,22), ylab="CO2 Emissions",main="VAR - 20 Year Forecast")
points(seq(37,56,1),varfore20,type="b",pch=15)

#rolling window ASE
#R_W_ASE(co2_total$CO2, trainingSize=20, horizon=5, s=0, d=1, thetas=c(-0.5532092), phis=0)

#rolling window ASE
R_W_ASE_VAR(alldata$CO2, trainingSize=20, horizon=5, model=varmodel)

# "The Summary Statistics for the Rolling Window ASE Are:"
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.4973  2.0379  7.6678  8.3183 14.6984 17.3606 
# "The Rolling Window ASE is:  8.31825144249936"



```

# Univariate MLP Model: 
```{r}
co2_all = ts(alldata$CO2[1:36], start =c(1981,1),frequency =1)
co2_5 = ts(alldata$CO2[1:31], start =c(1981,1),frequency =1)
co2_20 = ts(alldata$CO2[1:16], start =c(1981,1),frequency =1)
##-----------------------------------------------
##        -- Checking it with Mean
##-----------------------------------------------
#Using data minus 5 years
set.seed(2)
fit.mlp = mlp(co2_5,reps = 50,comb="mean")
fit.mlp 
plot(fit.mlp)

#Forecasting out 5 years ahead
fore.mlp = forecast(fit.mlp,h=5)
plot(fore.mlp, ylab="CO2 Emissions",main="MLP - 5 Year Forecast")

# 5yr ASE
mean((fore.mlp$mean - co2_all[32:36])^2)
##-----------------------------------------------
#Using data minus 20 years
set.seed(2)
fit.mlp2 = mlp(co2_20,reps = 50,comb="mean")
fit.mlp2 
plot(fit.mlp2)

#Forecasting out 20 years ahead
fore.mlp2 = forecast(fit.mlp,h=20)
plot(fore.mlp2, ylab="CO2 Emissions",main="MLP - 20 Year Forecast")

# 20 yr ASE
mean((fore.mlp2$mean - co2_all[17:36])^2)
##-----------------------------------------------
##        -- Checking it with Median 
##-----------------------------------------------
#Using data minus 5 years
set.seed(2)
fit.mlp = mlp(co2_5,reps = 50,comb="median")
fit.mlp 
plot(fit.mlp)

#Forecasting out 5 years ahead
fore.mlp = forecast(fit.mlp,h=5)
plot(fore.mlp, ylab="CO2 Emissions",main="MLP - 5 Year Forecast")

# 5yr ASE
mlpase5 = mean((fore.mlp$mean - co2_all[32:36])^2)
print(paste0("The 5 year ASE for the MLP model is: ", mlpase5))
##-----------------------------------------------
#Using data minus 20 years
set.seed(2)
fit.mlp2 = mlp(co2_20,reps = 50,comb="median")
fit.mlp2 
plot(fit.mlp2)

#Forecasting out 20 years ahead
fore.mlp2 = forecast(fit.mlp,h=20)
plot(fore.mlp2, ylab="CO2 Emissions",main="MLP - 20 Year Forecast")

# 20 yr ASE
mlpase20 = mean((fore.mlp2$mean - co2_all[17:36])^2)
print(paste0("The 20 year ASE for the MLP model is: ", mlpase20))
```
#Ensemble forecast
A forecast was developed using an averaging ensemble of two models: the VAR model and the ARMA model.
```{r}
#Ensemble ASE =======================
#average the ASE forecasts from the ARMA and VAR models
Ensase5 = (CO2_f5$f + varase5) / 2
Ensase20 = (CO2_f20$f + varase20) / 2

#Use the ASE helper function to get the ASE
StaticASE(series=alldata$CO2, Ensase5, n.ahead=5)
#"5 year ASE: 0.784638702011292"
StaticASE(series=alldata$CO2, Ensase20, n.ahead=20)
#"20 year ASE: 2.64724674114795"

#ASE Plotting =====================
#Plot the 5-year ensemble forecast
plot(seq(1,36,1),alldata$CO2,type="b", xlim=c(1,41), ylab="CO2 Emissions",main="Ensemble - 5 Year ASE")
points(seq(32,36,1),Ensase5,type="b",pch=15)

#Plot the 20-year ensemble forecast
plot(seq(1,36,1),alldata$CO2,type="b", xlim=c(1,56), ylab="CO2 Emissions",main="Ensemble - 20 Year ASE")
points(seq(17,36,1),Ensase20,type="b",pch=15)


#Ensemble Forecasting ================
#average the forecasts from the ARMA and VAR models
EnsFore5 = (armafore5$f + varfore5) / 2
EnsFore20 = (armafore20$f + varfore20) / 2

#Plotting =====================
#Plot the 5-year ensemble forecast
plot(seq(1,36,1),alldata$CO2,type="b", xlim=c(1,41), ylab="CO2 Emissions",main="Ensemble - 5 Year Forecast")
points(seq(37,41,1),EnsFore5,type="b",pch=15)

#Plot the 20-year ensemble forecast
plot(seq(1,36,1),alldata$CO2,type="b", xlim=c(1,56), ylab="CO2 Emissions",main="Ensemble - 20 Year Forecast")
points(seq(37,56,1),EnsFore20,type="b",pch=15)

#Rolling Window ASE ==================
#Rolling Window ASE for the Ensemble Model
R_W_ASE_ENS(alldata$CO2, trainingSize=20, horizon=5, varmodel=varmodel, phis=0.9261945, thetas=-0.5741199)

# "The Summary Statistics for the Rolling Window ASE Are:"
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.1713  0.4341  1.3144  1.7516  3.2597  4.6931 
# "The Rolling Window ASE is:  1.75163857109927"
```


